Comparing the perfomance impacts of narrow and extended feature spaces for machine learning bots in the game of Schnapsen

### 2018-02-01 Text/paper.tex
Final version of the paper and added inverse kbbot (play expensive)

\title{Comparing the perfomance impacts of narrow and extended feature spaces for machine learning bots in the game of Schnapsen}
\date{January 25, 2018}
\author{Lucas Faijdherbe, Daan Schrage and Ruben van der Ham\\\\
	\small Intelligent Systems\\
	\small Vrije Universiteit Amsterdam}

@@

\clearpage
\section{Research Question}
In our research, we want to find \textit{if extending the feature set will improve the machine learning bot's performance the most in the game of Schnapsen}. With improve the most, we mean as seen from the baseline. There is already a machine learning bot present, which will act as this base line. This bot has, what we will call the \textit{default feature space}- provided by the framework.  \\

@@

\item[ml\_combined] \hfill
\begin{itemize}
\item Perspective information
\item Player 1's point count
\item Player 1's point count squared
\item Player 1's point count cubed
\item Player 2's point count
\item Player 2's point count squared
\item Player 2's point count cubed
\item Player 1's pending points
\item Player 1's pending points squared
\item Player 1's pending points cubed
\item Player 2's pending points
\item Player 2's pending points squared
\item Player 2's pending points cubed
\item Difference between player 1 and player 2's point count
\item Difference between player 1 and player 2's pending point count
\item The trump suit
\item Trump card ratio (Number of trump cards in hand / total number of cards in hand * 100)
\item High-ranking card ratio (Number of 10's and A's in hand / total number of cards in hand * 100)
\item The phase
\item Number of trump cards in hand in phase 2
\item The stock size
\item The leader
\item Whose turn it is currently
\item The played card of the opponent
\item Number of cards in same suit as opponent's played suit in hand.
\item Number of points in hand squared ($card1^2$ + $card2^2$ ...)
\end{itemize}
\end{description}
\subsection{Worksheet I}
\begin{enumerate}

@@

\item We implemented the PlayCheap strategy:

\item Game output

\item 
\end{enumerate}

>>>>

\title{Comparing the perfomance impacts of narrow and extended feature spaces for machine learning bots in the game of Schnapsen}
\date{January 27, 2018}
\author{Lucas Faijdherbe, Daan Schrage and Ruben van der Ham\\\\
	\small Intelligent Systems\\
	\small Vrije Universiteit Amsterdam}

@@

\clearpage
\section{Research Question}
In our research, we want to find \textit{if extending the feature set will improve the machine learning bot's performance in the game of Schnapsen}. With improve the most, we mean as seen from the baseline. There is already a machine learning bot present, which will act as this base line. This bot has, what we will call the \textit{default feature space}- provided by the framework.  \\

@@

\item[ml\_combined] \hfill
\begin{itemize}
\item Perspective information
\item Player 1's point count
\item Player 1's point count squared
\item Player 1's point count cubed
\item Player 2's point count
\item Player 2's point count squared
\item Player 2's point count cubed
\item Player 1's pending points
\item Player 1's pending points squared
\item Player 1's pending points cubed
\item Player 2's pending points
\item Player 2's pending points squared
\item Player 2's pending points cubed
\item Difference between player 1 and player 2's point count
\item Difference between player 1 and player 2's pending point count
\item The trump suit
\item Trump card ratio (Number of trump cards in hand / total number of cards in hand * 100)
\item High-ranking card ratio (Number of 10's and A's in hand / total number of cards in hand * 100)
\item The phase
\item Number of trump cards in hand in phase 2
\item The stock size
\item The leader
\item Whose turn it is currently
\item The played card of the opponent
\item Number of cards in same suit as opponent's played suit in hand.
\item Number of points in hand squared ($card1^2$ + $card2^2$ ...)
\end{itemize}
\end{description}

\clearpage
\subsection{Worksheet I}
\begin{enumerate}

@@

\item We implemented the PlayCheap strategy:
\begin{lstlisting}[language=bash]
    kb.add_clause(~PC2, J2, Q2, K2)
    kb.add_clause(~J2, PC2)
    kb.add_clause(~Q2, PC2)
    kb.add_clause(~K2, PC2)

    kb.add_clause(~PC3, J3, Q3, K3)
    kb.add_clause(~J3, PC3)
    kb.add_clause(~Q3, PC3)
    kb.add_clause(~K3, PC3)

    kb.add_clause(~PC4, J4, Q4, K4)
    kb.add_clause(~J4, PC4)
    kb.add_clause(~Q4, PC4)
    kb.add_clause(~K4, PC4)

    kb.add_clause(~PC7, J7, Q7, K7)
    kb.add_clause(~J7, PC7)
    kb.add_clause(~Q7, PC7)
    kb.add_clause(~K7, PC7)

    kb.add_clause(~PC8, J8, Q8, K8)
    kb.add_clause(~J8, PC8)
    kb.add_clause(~Q8, PC8)
    kb.add_clause(~K8, PC8)

    kb.add_clause(~PC9, J9, Q9, K9)
    kb.add_clause(~J9, PC9)
    kb.add_clause(~Q9, PC9)
    kb.add_clause(~K9, PC9)

    kb.add_clause(~PC12, J12, Q12, K12)
    kb.add_clause(~J12, PC12)
    kb.add_clause(~Q12, PC12)
    kb.add_clause(~K12, PC12)

    kb.add_clause(~PC13, J13, Q13, K13)
    kb.add_clause(~J13, PC13)
    kb.add_clause(~Q13, PC13)
    kb.add_clause(~K13, PC13)

    kb.add_clause(~PC14, J14, Q14, K14)
    kb.add_clause(~J14, PC14)
    kb.add_clause(~Q14, PC14)
    kb.add_clause(~K14, PC14)

    kb.add_clause(~PC17, J17, Q17, K17)
    kb.add_clause(~J17, PC17)
    kb.add_clause(~Q17, PC17)
    kb.add_clause(~K17, PC17)

    kb.add_clause(~PC18, J18, Q18, K18)
    kb.add_clause(~J18, PC18)
    kb.add_clause(~Q18, PC18)
    kb.add_clause(~K18, PC18)

    kb.add_clause(~PC19, J19, Q19, K19)
    kb.add_clause(~J19, PC19)
    kb.add_clause(~Q19, PC19)
    kb.add_clause(~K19, PC19)
\end{lstlisting}

\item Game output:
\begin{lstlisting}[language=bash]
   Start state: The game is in phase: 1
Player 1's points: 0, pending: 0
Player 2's points: 0, pending: 0
The trump suit is: H
Player 1's hand: 10C QC 10D KD JH
Player 2's hand: AC KC AD QH 10S
There are 10 cards in the stock

player1: <bots.kbbot.kbbot.Bot instance at 0x7f1506109998>
player2: <bots.rand.rand.Bot instance at 0x7f1506109ef0>
*   Player 1 plays: JH
The game is in phase: 1
Player 1's points: 0, pending: 0
Player 2's points: 0, pending: 0
The trump suit is: H
Player 1's hand: 10C QC 10D KD JH
Player 2's hand: AC KC AD QH 10S
There are 10 cards in the stock
Player 1 has played card: J of H

*   Player 2 plays: 10S
The game is in phase: 1
Player 1's points: 12, pending: 0
Player 2's points: 0, pending: 0
The trump suit is: H
Player 1's hand: 10C QC 10D KD QD
Player 2's hand: AC KC AD QH AS
There are 8 cards in the stock

*   Player 1 plays: QD
*   Player 1 melds a marriage between QD and KD
The game is in phase: 1
Player 1's points: 12, pending: 20
Player 2's points: 0, pending: 0
The trump suit is: H
Player 1's hand: 10C QC 10D KD QD
Player 2's hand: AC KC AD QH AS
There are 8 cards in the stock
Player 1 has played card: Q of D

*   Player 2 plays: KC
The game is in phase: 1
Player 1's points: 39, pending: 0
Player 2's points: 0, pending: 0
The trump suit is: H
Player 1's hand: 10C QC 10D KD AH
Player 2's hand: AC AD QH AS KS
There are 6 cards in the stock

*   Player 1 plays: KD
The game is in phase: 1
Player 1's points: 39, pending: 0
Player 2's points: 0, pending: 0
The trump suit is: H
Player 1's hand: 10C QC 10D KD AH
Player 2's hand: AC AD QH AS KS
There are 6 cards in the stock
Player 1 has played card: K of D

*   Player 2 plays: AS
The game is in phase: 1
Player 1's points: 54, pending: 0
Player 2's points: 0, pending: 0
The trump suit is: H
Player 1's hand: 10C QC JC 10D AH
Player 2's hand: AC AD KH QH KS
There are 4 cards in the stock

*   Player 1 plays: JC
The game is in phase: 1
Player 1's points: 54, pending: 0
Player 2's points: 0, pending: 0
The trump suit is: H
Player 1's hand: 10C QC JC 10D AH
Player 2's hand: AC AD KH QH KS
There are 4 cards in the stock
Player 1 has played card: J of C

*   Player 2 plays: QH
The game is in phase: 1
Player 1's points: 54, pending: 0
Player 2's points: 5, pending: 0
The trump suit is: H
Player 1's hand: 10C QC 10D JD AH
Player 2's hand: AC AD KH KS QS
There are 2 cards in the stock

*   Player 2 plays: QS
The game is in phase: 1
Player 1's points: 54, pending: 0
Player 2's points: 5, pending: 0
The trump suit is: H
Player 1's hand: 10C QC 10D JD AH
Player 2's hand: AC AD KH KS QS
There are 2 cards in the stock
Player 2 has played card: Q of S

*   Player 1 plays: QC
The game is in phase: 2
Player 1's points: 54, pending: 0
Player 2's points: 11, pending: 0
The trump suit is: H
Player 1's hand: 10C 10D JD AH 10H
Player 2's hand: AC AD KH KS JS
There are 0 cards in the stock

*   Player 2 plays: KS
The game is in phase: 2
Player 1's points: 54, pending: 0
Player 2's points: 11, pending: 0
The trump suit is: H
Player 1's hand: 10C 10D JD AH 10H
Player 2's hand: AC AD KH KS JS
There are 0 cards in the stock
Player 2 has played card: K of S

*   Player 1 plays: 10H
The game is in phase: 2
Player 1's points: 68, pending: 0
Player 2's points: 11, pending: 0
The trump suit is: H
Player 1's hand: 10C 10D JD AH
Player 2's hand: AC AD KH JS
There are 0 cards in the stock

Game finished. Player 1 has won, receiving 2 points.
\end{lstlisting}
\item Game output of tournament kbbot with 'cheap strategy' and kbbot inverse with 'play expensive' strategy:
\begin{lstlisting}[language=bash]
Playing 120 games:
Results:
    bot <bots.kbbotinverse.kbbotinverse.Bot .......>: 64 wins
    bot <bots.kbbot.kbbot.Bot ......>: 56 wins
\end{lstlisting}
\end{enumerate}
